---
title: "Neff_CodingChallenge4"
author: "AJ Neff"
date: "2025-02-27"
output:
  md_document:
    variant: gfm
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
    toc_float: true
  word_document: null
---

The data is accessible via the link in this citation:
Noel, Z.A., Roze, L.V., Breunig, M., Trail, F. 2022. Endophytic fungi as promising biocontrol agent to protect wheat from Fusarium graminearum head blight. Plant Disease. 
[Link text](https://doi.org/10.1094/PDIS-06-21-1253-RE)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
a. YAML header includes title, author, date, output
b. Literate programming is a mixture of input code and outputs of code that allow for fully reproducible documents.

# Question 2

```{r}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
library(readr)
MycotoxinData <- read.csv("MycotoxinData.csv",na.strings = "na")
View(MycotoxinData)
```

```{r}
DON_data <- read.csv("MycotoxinData.csv", na.strings = "na")
str(DON_data)
```

## CC3 Q1
```{r}
library(ggplot2)
DON_plot <- ggplot(MycotoxinData, aes(x = Treatment, y = DON, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab("DON (ppm)") +
  geom_jitter(alpha = 0.6, pch = 21, color = "black", position = position_jitterdodge()) +
  scale_fill_manual(values = c(cbbPalette[[2]], cbbPalette[[8]])) +
  facet_wrap(~Cultivar) + 
  theme_classic()
DON_plot
recordPlot()
```

## CC3 Q2
```{r}
MycotoxinData$Treatment <- factor(MycotoxinData$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))
```

## CC3 Q3
```{r}
DON15 <- ggplot (MycotoxinData, aes(x = Treatment, y = X15ADON, fill = Cultivar)) +
  geom_boxplot(outliers = F) +
  xlab("") +
  ylab("DON (ppm)") +
  geom_jitter(alpha = 0.6, pch = 21, color = "black", position = position_jitterdodge()) +
  scale_fill_manual(values = c(cbbPalette[[2]], cbbPalette[[8]])) +
  facet_wrap(~Cultivar) + 
  theme_classic()
DON15
recordPlot()
```

```{r}
DONSeed <- ggplot (MycotoxinData, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar)) +
  geom_boxplot(outliers = F) +
  xlab("") +
  ylab("DON (ppm)") +
  geom_jitter(alpha = 0.6, pch = 21, color = "black", position = position_jitterdodge()) +
  scale_fill_manual(values = c(cbbPalette[[2]], cbbPalette[[8]])) +
  facet_wrap(~Cultivar) + 
  theme_classic()
DONSeed
recordPlot()
```

## CC3 Q4
```{r}
library(tidyverse)
library(ggpubr)
library(ggrepel)
```

```{r}
ggarrange(DON_plot, DON15, DONSeed, labels = "auto", ncol = 3, nrow = 1, common.legend = T)
```

## CC3 Q5
```{r}
stats_DON_plot <- DON_plot +
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
stats_DON_plot
recordPlot()
```

```{r}
stats_DON15 <- DON15 +
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
stats_DON15
recordPlot()
```

```{r}
stats_DONSeed <- DONSeed +
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
stats_DONSeed
recordPlot()
```

## Combined
```{r}
ggarrange(stats_DON_plot, stats_DON15, stats_DONSeed, labels = "auto", ncol = 3, nrow = 1, common.legend = T)
```

# Question 3

