---
title: "CodingChallenge6"
author: "AJ Neff"
date: "2025-03-27"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
The main point is simplify your code so that it's easy to use repeatably without error.

# Question 2
A function must first be assigned a name and given the beginning of the input code such as the syntax below:
    function_name <- function(variable) {
      body
      return(output)
    }.
You then add the code in the body and a return for what you want the output to do.  

# Question 3
```{r}
library(tidyverse)
library(ggplot2)
Cities <- read.csv("Cities.csv")
```

# Question 4
```{r}
# convert to radians
dist_bet <- function(lat1, lon1, lat2, lon2){
  rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180
# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 
# Earth's radius in kilometers
earth_radius <- 6378137
# Calculate the distance
distance_km <- (earth_radius * c)/1000
return(distance_km)
}
```

# Question 5

```{r}
Cities_subset <- subset(Cities, city %in% c("New York", "Auburn"), select = c(long, lat))
print(Cities_subset)
nyc_auburn <- dist_bet(40.6943, -73.9249, 32.6087, -85.4903)
print(nyc_auburn)
```

# Question 6

```{r}
dist_all <- NULL
for (i in 1:nrow(Cities)){
  distance_i <- dist_bet(Cities$lat[i], Cities$long[i], 32.6087, -85.4903)
  dist_all <- rbind.data.frame(dist_all, distance_i)
}

print(dist_all)
```

# Question 7

I forgot how to link it. 